Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2025-01-22 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
sweat test data problems

```{r data cleaning}
#| message: false
# Import data
# source("/Users/timvigers/Library/CloudStorage/Dropbox/GitHub/CF/Edith Zemanick/Sweat Variability/data_import.r")
load("./Data_Cleaned/chco_data.RData")
# Exclude people without demographics
chco <- chco %>% filter(demog_avail == "True")
# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) |
  diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
chco <- chco[!(chco$patient_id %in% exclude), ]
chco$Diagnosis <- diagnoses$Diagnosis[match(chco$patient_id, diagnoses$PatientID)]
diagnosis_check <- chco %>% select(patient_id, Diagnosis, diagnosis)
# Import NJH data, remove those without demographics
njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx")
# Filter and select similar to CHCO
njh <- njh %>% filter(`Demographics Available`)
# Import detailed diagnoses from Katie
diagnoses <- read.csv("./Data_Cleaned/SweatVariability2025_CFgenotypes.csv",
  na.strings = c("", ".")
)
# Merge NJH and diagnoses
diagnoses <- diagnoses %>% select(Patient.ID, Diagnosis...DataScout)
colnames(diagnoses) <- c("PatientID", "Diagnosis")
# Less detailed diagnoses
diagnoses$Diagnosis <- factor(diagnoses$Diagnosis,
  levels = c(
    "Abnormal sweat choride test without diagnosis of cystic fibrosis",
    "Abnormal sweat test",
    "Alpha-1-antitrypsin deficiency",
    "Alpha-1-antitrypsin deficiency carrier",
    "Alpha-1-antitrypsin deficiency, Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "CF", "CF carrier, Alpha-1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency carrier",
    "CF, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency carrier",
    "Cystic fibrosis screening",
    "Cystic fibrosis transmembrane conductance regulator (CFTR)-related disorder",
    "Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "Primary ciliary dyskinesia",
    "Primary Ciliary dyskinesia, Cystic Fibrosis carrier"
  ),
  labels = c(
    "Non-CF", "Non-CF", "Non-CF", "CF Carrier", "Non-CF", "Cystic Fibrosis",
    "CF Carrier", "Cystic Fibrosis", "Cystic Fibrosis", "Cystic Fibrosis",
    "CF Carrier", "CF Carrier", "CF Carrier", "Non-CF", "CFTR-RD or CRMS",
    "Non-CF", "Non-CF", "CF Carrier"
  )
)
# Patient ids may not match so keep CHCO and NJH separate and check names after
```

>>>>

```{r data cleaning}
#| message: false
# Import data
chco <- read.csv("./Data_Raw/SweatVariability2025_DATA_LABELS_2025-01-21_1211.csv",
  na.strings = ""
)
# Format dates
chco$Birth.Date <- as.Date(chco$Birth.Date, "%Y-%m-%d")
chco$Sweat.test.date <- as.Date(chco$Sweat.test.date, "%Y-%m-%d")
# Exclude people without demographics
chco <- chco %>% filter(Demographics.Available == "True")
# Fix diagnoses/exclusion data for CHCO (based on Angela's review)
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude) |
  diagnoses$Code == "Exclude"]
exclude <- exclude[!is.na(exclude)]
chco <- chco[!(chco$Patient.ID %in% exclude), ]
chco$New.Diagnosis <- diagnoses$Diagnosis[match(chco$Patient.ID, diagnoses$PatientID)]
diagnosis_check <- chco %>% select(Patient.ID, Diagnosis, New.Diagnosis)
# Import NJH data, remove those without demographics
njh <- read_excel("./Data_Raw/SweatVariability2025_CFgenotypes 2024-08-21.xlsx", .name_repair = "universal")
# Filter and select similar to CHCO
njh <- njh %>% filter(Demographics.Available)
# Convert dates
njh$Birth.Date <- as.Date(njh$Birth.Date, origin = "1899-12-30")
njh$Sweat.test.date <- as.Date(njh$Sweat.test.date, origin = "1899-12-30")
# Import detailed diagnoses from Katie
diagnoses <- read.csv("./Data_Cleaned/SweatVariability2025_CFgenotypes.csv",
  na.strings = c("", ".")
)
# Merge NJH and diagnoses
diagnoses <- diagnoses %>% select(Patient.ID, Diagnosis...DataScout)
colnames(diagnoses) <- c("PatientID", "Diagnosis")
# Less detailed diagnoses
diagnoses$Diagnosis <- factor(diagnoses$Diagnosis,
  levels = c(
    "Abnormal sweat choride test without diagnosis of cystic fibrosis",
    "Abnormal sweat test",
    "Alpha-1-antitrypsin deficiency",
    "Alpha-1-antitrypsin deficiency carrier",
    "Alpha-1-antitrypsin deficiency, Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "CF", "CF carrier, Alpha-1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency",
    "CF, alpha 1-antitrypsin deficiency carrier",
    "CF, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency",
    "Cystic fibrosis carrier, Alpha-1-antitrypsin deficiency carrier",
    "Cystic fibrosis screening",
    "Cystic fibrosis transmembrane conductance regulator (CFTR)-related disorder",
    "Positive cystic fibrosis sweat test without diagnosis of cystic fibrosis",
    "Primary ciliary dyskinesia",
    "Primary Ciliary dyskinesia, Cystic Fibrosis carrier"
  ),
  labels = c(
    "Non-CF", "Non-CF", "Non-CF", "CF Carrier", "Non-CF", "Cystic Fibrosis",
    "CF Carrier", "Cystic Fibrosis", "Cystic Fibrosis", "Cystic Fibrosis",
    "CF Carrier", "CF Carrier", "CF Carrier", "Non-CF", "CFTR-RD or CRMS",
    "Non-CF", "Non-CF", "CF Carrier"
  )
)
# Patient ids may not match so keep CHCO and NJH separate and check names after
njh$New.Diagnosis <- diagnoses$Diagnosis[match(njh$Patient.ID, diagnoses$PatientID)]
diagnosis_check_njh <- njh %>% select(Patient.ID, Diagnosis, New.Diagnosis)
# Remove unnecessary columns and combine
chco <- chco %>%
  mutate(Dataset = "CHCO") %>%
  select(
    Dataset, Record.ID, Patient.ID, Patient.Name, Birth.Date, Diagnosis,
    New.Diagnosis, Genotype.1, Genotype.2, Sweat.test.date,
    Age.at.test..in.years., Site.of.test:Sweat.chloride.right.arm
  )
njh <- njh %>%
  mutate(Dataset = "NJH") %>%
  select(
    Dataset, Record.ID, Patient.ID, Patient.Name, Birth.Date, Diagnosis,
    New.Diagnosis, Genotype.1, Genotype.2, Sweat.test.date,
    Age.at.test..in.years., Site.of.test:Sweat.chloride.right.arm
  )
df <- rbind(chco, njh)
df <- df %>% arrange(Patient.ID, Sweat.test.date, Dataset)
# Save
write.csv(df, file = "./Data_Cleaned/combined_data.csv", row.names = F, na = "")
```
